Early Observations of the Type Ia Supernova \abc:\\ Evidence for Strong Ejecta Mixing and/or Interaction with Diffuse Material

### 2017-08-17 doc/ms.tex
Update text to describe new companion-interaction fig

The detection of excess emission due to the collision of the SN ejecta 
with a non-degenerate companion requires a favorable orbital 
alignment relative to the line of sight. Thus, from geometric 
considerations alone the probability of detecting ejecta-companion 
interaction is low, $\sim$10\%. \citet{2010ApJ...708.1025K} calculates that the 
collision of SN ejecta with a companion  
generates thermal emission with a spectrum that peaks in the 
UV. The resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN-companion signature in the early
light curve of \abc, we employ the \citet{2010ApJ...708.1025K}
model and assume canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity,
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity, 
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected 
$g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the 
distance of \abc\ behind a total reddening of $E(B-V) = 0.08 \, \mathrm{mag}$ using the parameterized equations in 
\citet{2012ApJ...749...18B}. If we assume the binary is aligned with the optimal orientation relative to the line of sight, a binary separation of $a \approx 2 \times 10^{13}\, \mathrm{cm}$ is needed to explain the initial detection of \abc, as shown in Figure~\ref{fig:SN-companion}. This model does not, however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 1 \, \mathrm{d}$ (though it is possible that the SN photosphere dominates the companion-interaction signature at this phase). Furthermore, this model significantly overpredicts the observed flux in the \textit{Swift}/$UVW1$ filter $\sim$3.1~d after $t_0$. As such, we do not favor such a model as the explanation for the early emission from \abc.

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a = 10^{14} \, \mathrm{cm}$ provides a good match to the observations $\sim$1--3~d after $t_0$ if $t_\mathrm{exp} \approx t_0 + 0.3 \; \mathrm{d}$. Like the $a = 2 \times 10^{13} \, \mathrm{cm}$ model, this model significantly overpredicts the observed UV flux from \abc. Additionally, this model cannot account for the brightness of \abc\ at the epoch of discovery. The challenges associated with each of the previous two models lead us to conclude that the early evolution of \abc\ cannot be explained via the interaction of the SN ejecta with a red-giant companion.\footnote{The \citet{2010ApJ...708.1025K} model assumes Roche-lobe overflow, meaning that $a > 10^{13} \, \mathrm{cm}$ implies a companion with $R_\ast \gtrsim 100 \, R_\odot$.} We cannot, however, exclude the presence of a red-giant companion as our calculations have assumed that the binary is aligned with the optimal geometry relative to the line of sight. If the geometry is not favorable, then it is possible that signatures from interaction with a companion are not visible.

 % Our early \textit{Swift} observations support this conclusion, as the UV $-$ optical colors are very red, whereas iPTF\,14atg exhibited much bluer UV $-$ optical colors at the same epoch \citep{2015Natur.521..328C}.

@@

Finally, we examine the minimum binary separation capable of explaining the observed brightness at the epoch of discovery. Figure~\ref{fig:SN-companion} shows that models with $a \approx 8 \times 10^{11} \, \mathrm{cm}$ peak at $\sim$21.5~mag in the $g_\mathrm{PTF}$-band, provided that $t_\mathrm{exp} \approx t_0 - 1 \; \mathrm{d}$. To explain just the epoch of discovery, models with larger $a$ would require even earlier explosion times, which is not compatible with our non-detections the night before discovery, while models with smaller $a$ are not bright enough. While it is otherwise compatible with the observations, we do not favor the $a \approx 8 \times 10^{11} \, \mathrm{cm}$ model as the explanation for the early flux from \abc\ because it does not explain the 2 mag rise in the $\sim$24 hr after discovery.

>>>>

The detection of emission from the collision of the SN ejecta 
with a non-degenerate companion requires a favorable orbital 
alignment relative to the line of sight. Thus, from geometric 
considerations alone the probability of detecting ejecta-companion 
interaction is low, $\sim$10\%. \citet{2010ApJ...708.1025K} 
calculates that the collision of SN ejecta with a companion  
generates thermal emission with a spectrum that peaks in the 
UV. The resulting \textit{g}-band emission is expected to be weak.

@@

To examine the possibility of a SN-companion signature in the early
light curve of \abc, we employ the \citet{2010ApJ...708.1025K}
model and assume canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity, $10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity, $0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected $g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the distance of \abc\ behind a total reddening of $E(B-V) = 0.08 \, \mathrm{mag}$ using the parameterized equations in \citet{2012ApJ...749...18B}. If we assume the binary is aligned with the optimal orientation relative to the line of sight, a binary separation of $a \approx 3 \times 10^{11}\, \mathrm{cm}$ is needed to explain the initial detection of \abc, as shown in Figure~\ref{fig:SN-companion}.  The minimum binary separation capable of explaining the observed brightness at the epoch of discovery is $a \approx 10^{11} \, \mathrm{cm}$. Figure~\ref{fig:SN-companion} shows that such models peak at $g_\mathrm{PTF} \approx 21.5 \, \mathrm{mag}$, provided that $t_\mathrm{exp} \approx t_0 - 0.3 \; \mathrm{d}$. These models do not, however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 0.5 \, \mathrm{d}$ (though it is possible that the SN photosphere dominates the companion-interaction signature at this phase). While they are otherwise compatible with the observations, we do not favor the above models as the explanation for the early flux from \abc\ because they do not explain the 2 mag rise in the $\sim$24 hr after discovery.

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a \approx 18 \times 10^{11} \, \mathrm{cm}$ provides a good match to the initial $g_\mathrm{PTF}$ rise, if $t_\mathrm{exp} \approx t_0 + 0.1 \; \mathrm{d}$. Models with $a \gtrsim 10^{12} \, \mathrm{cm}$, which can explain the initial $g_\mathrm{PTF}$ rise, signiticantly overpredict the observed UV flux, however. There is no choice of $a$ capable of replicating the early rise of \abc\ without also significantly overpredicting the observed UV flux.

The challenges associated with each of the previously considered models lead us to conclude that the early evolution of \abc\ cannot be explained via the interaction of the SN ejecta with a companion. We cannot, however, exclude the presence of a red-giant, or other non-degenerate, companion as our calculations have assumed that the binary is aligned with the optimal geometry relative to the line of sight. If the geometry is not favorable, then it is possible that signatures from interaction with a companion are not visible.
